### 2020-06-03 report/sections/server/node/alt.tex
Report: progress on chapter 4

The main alternative EFSM encoding would be similar to those
presented in \cite{Hybrid2016}, which encodes
each EFSM state as a separate class and expose communication APIs
(e.g. send and/or receive) that respect the permitted transitions at
that state.

\begin{lstlisting}[language=scribble]
global protocol OneAdder(role Client, role Svr) {
	NUM1(number) from Client to Svr;
	NUM2(number) from Client to Svr;
	SUM(number)  from Svr to Client;
}
\end{lstlisting}

\begin{figure}[!h]
\centering
\begin{subfigure}{0.49\textwidth}
\begin{lstlisting}[language=javascript,tabsize=2]
const logic = async (init) => {
	const [x, num2] = 
		await init.receive();
 	const [y, sum] = 
 		await num2.receive();

 	// init.receive(); (*@\label{line:nodebreaklinear}@*)
	return sum.send(x + y);
}
\end{lstlisting}
\caption{Alternative Version}
%\label{...}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
\begin{lstlisting}[language=javascript,tabsize=2]
const logic = new S4({
	NUM1: (x) => new S6({
		NUM2: (y) => ([
			Labels.S7.SUM,
			[x + y],
			new S5(),
		]),
	}),
});

\end{lstlisting}
\caption{Current Version}
%\label{...}
\end{subfigure}
\captionof{figure}{\dots}
%\label{...}
\end{figure}

>>>>

The main alternative EFSM encoding would be similar to those
presented in \cite{Hybrid2016}, which encodes
each EFSM state in its own class 
and expose communication APIs (e.g. send and/or receive) 
that respect the permitted state transitions as instance methods.

\begin{figure}[!h]
\centering
\begin{subfigure}{\textwidth}
\begin{lstlisting}[language=scribble]
global protocol OneAdder(role Client, role Svr) {
	NUM1(number) from Client to Svr;
	NUM2(number) from Client to Svr;
	SUM(number)  from Svr to Client;
}
\end{lstlisting}
\caption{The \tprotocol{OneAdder} Protocol}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
\begin{lstlisting}[language=javascript,tabsize=2]
const logic = async (init) => {
	const [x, num2] = 
		await init.receive();
 	const [y, sum] = 
 		await num2.receive();

 	// init.receive(); (*@\label{line:nodebreaklinear}@*)
	return sum.send(x + y);
}
\end{lstlisting}
\caption{Alternative API Design}
%\label{...}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
\begin{lstlisting}[language=javascript,tabsize=2]
const logic = new S4({
	NUM1: (x) => new S6({
		NUM2: (y) => ([
			Labels.S7.SUM,
			[x + y],
			new S5(),
		]),
	}),
});

\end{lstlisting}
\caption{Current API Design}
%\label{...}
\end{subfigure}
\captionof{figure}{\dots}
%\label{...}
\end{figure}

### 2020-06-03 report/sections/server/node/efsm.tex
Report: progress on chapter 4

\subparagraph{Messages}

@@

By expressing the payload type as a \textit{tuple}\footnote{
In TypeScript, a tuple is an array with fixed size
and known types for elements at each position.
},
we easily generalise our type definition to polyadic payloads.

@@

We collect the APIs that the developer needs to implement
under the \texttt{Handler} namespace. 
As a design choice, we \textit{do not} generate handlers for
terminal states, because the semantics of inactivity mean
there is nothing to handle.
%Hence, all generated handler APIs are for non-terminal states,
%and will hold some reference to the successor state encoding.
%The reader will notice that, in the listings below, the successor
%state is stated to be under the 
We outline how the generated API differs between sending
and receiving states.

\subparagraph{Send}
We model selections using a \textit{union type} to
encapsulate the possible send actions, as shown in 
\cref{lst:addersvrsendhandler}.
Each send action is encoded as a tuple of
the label, the payload, and the successor state encoding.
We see some benefits from defining Message Types as interfaces:
with deterministic properties, we can index into the message type
to access the property's type, 
e.g. \lstonelinejs{Message.S54THANKS['payload']} 
would resolve to
\lstonelinejs{[string]},
based on \cref{lst:addersvrmsg}.

@@

The interface values are functions parameterised by
the message payload, and must return the successor state encoding.
We see another benefit of defining the payload in Message Types
as a tuple: we can define the receive handler parameter
using the \textit{spread syntax}, which allows the tuple
expression to be expanded into a list of arguments. 
More concretely, as shown in \cref{lst:nodeefsmspread}
it allows the developer to pattern match on the
individual payload values (\cref{line:yesspread}) 
rather than defining their function to expect a tuple 
and manually destructing it (\cref{line:nospread}),
so the former is more intuitive.

@@

\subsection{Wrapping Handlers in ``Implementations''}
\label{subsection:nodeefsmimplementation}
\begin{itemize}
\item need a way for the runtime to distinguish between states (implementation)
\item attempt at using conditional types, but because they handle conditional union types in a distributed way (give example), cannot be statically typed
\item alternative -- use discriminated union by wrapping in an Implementation class
\item all non-terminal states will `return' the successor implementation -- very difficult to resolve in the type-checker in the runtime (give example of what it may look like) -- solve by giving each state implementation the `advance' function to respect the event-driven nature of everything, so they can call it on completion
\item visualise the interaction between runtime and states as a sequence diagram for message passing
\end{itemize}

>>>>

\subparagraph{Messages}
The handler APIs that we generate for developers
need to refer to the label identifier and payload type: 
we refer to this as the message structure, and encode this as a
Message Type. 
Each message is expressed as an interface with
properties for the label and payload.
These interfaces are grouped based on the EFSM state
they belong using \textit{union types}.
We illustrate this in \cref{lst:addersvrmsg}.
By expressing the payload type as a \textit{tuple}\footnote{
In TypeScript, a tuple is an array with fixed size
and known types for elements at each position.
},
we easily generalise our type definition to polyadic payloads.

@@

@@

We collect the APIs that the developer needs to implement
under the \texttt{Handler} namespace. 
As a design choice, we \textit{do not} generate handlers for
terminal states, because the semantics of inactivity mean
there is nothing to handle.
We introduce the generated handlers for sending and receiving states.
These are non-terminal states that will involve the encoding of its
successor state. The reader will notice that, in the listings below,
the successor state is stated to be under the
\texttt{Implementation} namespace: we explain in 
\cref{subsection:nodeefsmimplementation}, but for now,
it is sufficient to acknowledge that those refer to the encoding
of the successor state.

\subparagraph{Send}
We model selections using a union type to
encapsulate the possible send actions, as shown in 
\cref{lst:addersvrsendhandler}.
Each send action is encoded as a tuple of
the label, the payload, and the successor state encoding.
We see some benefits from defining Message Types as interfaces:
with deterministic properties, we can index into the message type
to access the property's type, 
e.g. \lstonelinejs{Message.S54THANKS['payload']} 
would resolve to
\lstonelinejs{[string]},
based on \cref{lst:addersvrmsg}.

@@

The interface values are functions parameterised by
the message payload, and must return the successor state encoding.
We see another benefit of defining the payload in Message Types
as a tuple: we can define the receive handler parameter
using the \textit{spread syntax}, which allows the tuple
expression to be expanded into a list of function arguments. 
More concretely, as shown in \cref{lst:nodeefsmspread},
it allows the developer to pattern match on the
individual payload values (\cref{line:yesspread}) 
rather than defining their function to expect a tuple 
and manually destructing it (\cref{line:nospread}),
so the former is more intuitive.

@@

\subsection{Wrapping Handlers in ``Implementations''}
\label{subsection:nodeefsmimplementation}

The behaviour of the runtime is dependent on the current state,
so it needs a way to distinguish between
all the different states -- one can think of this as implementing
the state transition function from the theory, which is analogue to 
overloading a \texttt{next()} method for each state.
Due to limitations in the TypeScript language, this would have to
be some sort of switch statement, with the \texttt{next()}
method parameterised by some base type assignable to all states.
Currently, the state is only determined by the handler
to be implemented by the developer, so the switch statement
and base type would have to be defined on the handler APIs.

Unfortunately, this is not practical. 
Handlers for send states are union types and
handlers for receive states are interfaces,
both of which are not supported 
by the \lstonelinejs{instanceof}
operator.

We attempt to address this by defining an enum of state identifiers
for each type of state (i.e. an enum for send states, 
an enum for receive states)
upon which to execute the EFSM, which solves the switch statement
problem.
Now, we are left with defining a mapping between the 
state identifier enum to the handler type. This construct would be
analogue to \textit{dependent types}, which again, is not a feature
of the TypeScript type system.

\begin{itemize}
\item need a way for the runtime to distinguish between states (implementation)
\item attempt at using conditional types, but because they handle conditional union types in a distributed way (give example), cannot be statically typed
\item alternative -- use discriminated union by wrapping in an Implementation class
\item all non-terminal states will `return' the successor implementation -- very difficult to resolve in the type-checker in the runtime (give example of what it may look like) -- solve by giving each state implementation the `advance' function to respect the event-driven nature of everything, so they can call it on completion
\item visualise the interaction between runtime and states as a sequence diagram for message passing
\end{itemize}
